# Activity Monitoring

### I will download the file and read in the data
```{r}
if (!file.exists("./Week2")) {
  dir.create("./Week2")
}
url = "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(url, destfile = "./Week2/Dataset.zip")
download <- date()
unzip(zipfile = "./Week2/Dataset.zip", exdir = "./Week2")
files <- list.files("./Week2", recursive = T)
activity <- read.csv("./Week2/activity.csv")
nwdate <- as.Date(activity$date)
steps <- activity$steps
interval <- activity$interval
```
### This is the histogram, mean, and median of the total number of steps per day

```{r}
totalSteps <- aggregate(steps~nwdate, data=activity, FUN=sum, na.rm=TRUE)
hist(totalSteps$steps, xlab="Steps", ylab="# of Days",col="blue",  breaks=10,main="Total # of Steps/Day")
print(paste("Mean: ", as.character(mean(totalSteps$steps))))
print(paste("Median: ", as.character(median(totalSteps$steps))))
```

### This is the time series plot of the average # of steps/day by interval
```{r}
stepsByInterval <- aggregate(steps~interval, activity, mean)
with(stepsByInterval, plot(interval, steps, xlab = "Interval", ylab = "Avg # of Steps", type = "l"))
```

### This is the 5-minute interval that contains the most steps
```{r}
maxstep <- stepsByInterval[which.max(stepsByInterval[,2]),1]
print(maxstep)
```

### I will fill in missing data from "steps" with the mean # of steps per interval
```{r}
head(activity)
stepNA <- is.na(activity[,1]) # All NAs are in the steps column
countStepNA <- length(stepNA[stepNA==TRUE])
intmean <- mean(stepsByInterval$steps)
nwactivity <- activity
nwactivity[stepNA,1] <- intmean
head(nwactivity)
```

### I use the mean of steps as the impute value for the missing step values and then create a new Histogram, mean, and median
```{r}
nwTotalSteps <- aggregate(steps ~ nwdate, data=nwactivity, FUN=sum, na.rm=TRUE)
hist(nwTotalSteps$steps, xlab="Steps",col="blue",  ylab="# of Days", breaks=10,main="New Total # of Steps/Day")
meanAfterImput <- mean(nwTotalSteps$steps)
print(meanAfterImput)
print(mean(totalSteps$steps), na.rm=TRUE)
medAfterImput <- median(nwTotalSteps$steps)
print(medAfterImput)
print(median(totalSteps$steps), na.rm=TRUE)
```

Imputing missing data has no impact on the mean, there is a minimal change in the median

### I create a mean variable for dayType (weekday, weekend) and interval and create a plot panel using lattice
```{r, message=FALSE}
library(lattice)
library(dplyr)
nwactivity$date <- as.Date(nwactivity$date)
nwactivity2 <- nwactivity %>% mutate(dayType = ifelse(weekdays(nwactivity$date)=="Saturday"|weekdays(nwactivity$date)=="Sunday","Weekend","Weekday"))
dayTypeAndIntMn <- nwactivity2 %>% group_by(dayType, interval) %>% summarise(dayTypeAndIntMn=mean(steps))
with(dayTypeAndIntMn, xyplot(dayTypeAndIntMn~interval|dayType, type="l", xlab = "Daily Intervals",
                             ylab = "Avg # of Steps", main = "Total Steps by DayType"))
```

Yes, there is a difference between the two plots.